---
title: "Assessment 1 – Data set"
author: "A1844790 - Dylan Rohan"
date: "04/11/2021"
output: word_document
---


```{r}
library(tidyverse)
library(dplyr)
library(stringr)
ashes <- read_csv("C:\\Users\\rohad\\OneDrive\\Documents\\Data science\\Data Taming, modelling and Vizalization_RStudio\\a1\\a1\\ashes.csv")
#double slashes for windows directory
ashes <- read_csv("ashes.csv")
#testing something mentioned in the discussion board, neat!
ashes
#Checking the table, currently a tibble of 27 x 13
unique(ashes$team)
#need to correct variable English to be England
unique(ashes$role)
#many duplicates under alternate variable names, eb. bat, batsman, batting
```

## Question One: Reading and Cleaning
##### 1.1
(Note: I finished the assignment very early, so I went back and put the code in for the way the course probably expected to see it as well.)

1.1)	For our analysis, the subjects are not the cricketers themselves, but each batting innings they participated in. In order to make the data tidy each subject needs its own row. Rearrange the data into a long format so that there is a row for each batter in each innings. Your new tibble should have 270 rows. 
[2 points] 

Each cell should represent only one measurement. Use str_match() to create new columns for each of the following for each player innings: 
*	the player’s batting number
*	their score
*	the number of balls they faced. 
[2 points] 


```{r}
colnames(ashes)
ashes_longform <- gather(ashes, key = "innings", value = "description", "Test 1, Innings 1" : "Test 5, Innings 2")
ashes_longform
#tibble now in long form, 270 x 5
ashes_innings_first <- ashes_longform[c(4, 1, 2, 3, 5)]
ashes_innings_first
#tibble now in long form with subject first

```

##### 1.2
1.2)	Recode the data to make it ‘tame’, that is:
*	ensure all categorical variables with a small number of levels are coded as factors, 
*	ensure all categorical variables with a large number of levels are coded as characters,
* ensure all quantitative variables are coded as integers or numeric, as appropriate. 
[3 points] 

```{r}
colnames(ashes)
ashes_longform <- gather(ashes, key = "innings", value = "description", "Test 1, Innings 1" : "Test 5, Innings 2")
ashes_longform
#tibble now in long form, 270 x 5
ashes_innings_first <- ashes_longform[c(4, 1, 2, 3, 5)]
ashes_innings_first
#tibble now in long form with subject first

```
##### 1.3
Clean the data; recode the factors using fct_recode() such that there are no typographical errors in the team names and player roles. 
[2 points] 


```{r}
order <- str_match(ashes_innings_first$description, "Batting at number ..")
with_order <- cbind(ashes_innings_first, order)
with_order
#Order now has its own column
runs <- str_match(with_order$description, "scored ....")
runs
with_runs <- cbind(with_order, runs)
#runs now has its own column
no._of_balls <- str_match(with_runs$description, "from ....")
no._of_balls
all_columns<- cbind(with_runs, no._of_balls)
#no.  of balls now has its own column
batting_order <- str_replace_all(all_columns$order, "[^0-9.-]", "")
runs_ <- str_replace_all(all_columns$runs, "[^0-9.-]", "")
balls_ <- str_replace_all(all_columns$no._of_balls, "[^0-9.-]", "")  
#Taking numerical values from strings
order1 <- tibble(batting_order)
runs1 <- tibble(runs_)
balls1 <- tibble(balls_)
#making data frames from those values
a1_o <- cbind(ashes_innings_first, order1)
a1_o_r <- cbind(a1_o, runs1)
a1_o_r_b<- cbind(a1_o_r, balls1)
#Order same, so binding columns
a1o_r_b <- a1_o_r_b$description <- NULL
a1_o_r_b <- a1_o_r_b %>%
  mutate_all(na_if, "")
#removing description column
a1_o_r_b
#now a tibble of 270 x 7 (removed description, but unused data is still accessible in "ashes_innings_first')
  

  #_______________________________________________________#
#WAS A KEEN BEEN, SO ALTERNATIVELY:

trial <- ashes_innings_first %>%
  mutate("runs_"=str_match(description,"from ....") , "batting_order" = str_match(description, "Batting at number .."), "balls_" = str_match(description, "scored ...."))
#description string broken into appropriate columns
trial <- trial %>%
  mutate("runs_" = str_replace_all(trial$runs_, "[^0-9.-]",""), "balls_"=str_replace_all(trial$balls_, "[^0-9.-]",""), "batting_order"=str_replace_all(trial$batting_order, "[^0-9.-]",""))
  mutate_all(trial, na_if, "") %/%
#Left the description column in here, but all is right with the world
trial
   # _______________________________________________________#
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
